### 2020-02-05 paper/Manuscript.Rmd
Writing 5-2-2020

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud, increasing awareness of questionable research practices, and the replication crisis. However, Open science is not merely a cure for this crisis - it is also an opportunity. Technological advances now enable us to perform reliable, cumulative, collaborative science more easily. Capitalizing on these advances has the potential to accellerate scientific progress. This paper introduces WORCS: Workflow for Open, Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It is easy to implement, and still ensures reproducibility in most circumstances. Adoption is further facilitated by the availability of a WORCS project template.

@@

One of the most impactful contributions to Open Science has been the development of the TOP-guidelines, and the effort to urge publishers to adopt them. WORCS defines the goals of Open Science in terms of the eight standards described in the TOP-guidelines: Comprehensive citation of literature, data, materials, and methods (1); data sharing (2), sharing code required to reproduce analyses (3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis (7); and replication of published results (8). WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations. Yet the value of grassroots change should not be underestimated. Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support. The goal of this paper is therefore to present a workflow that can help individual scientists to meet the principles of Open Science. This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist,

@@

<!-- 
* Grass-roots
* Open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction
-->

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science. Second, all of the aforementioned tools for an Open Science workflow are implemented in R. In general, many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Third, at the time of writing, R is the second-most cited statistical software package [@PopularityDataScience2012], following SPSS, which has no support for any of the tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format. Working with R is simplified immensely by using the Rstudio user interface for R. Rstudio integrates many of the tools used in WORCS in a user-friendly manner, and automates and streamlines many tedious or complicated aspects of working with R. This renders Rstudio a catch-all solution for generating open, reproducible code.

Another consideration is to develop a workflow that relies on text-based files only. These files can be read by machines and humans alike, without requiring special (licensed) software. Text-based files are also small, can be version-controlled with Git (binary files should not be version controlled), and can be perused online in repositories such as GitHub.

@@

- Use RStudio projects to bundle analyses and writing into a self-contained unit
- Use text-based files only
    * `.R` for analyses
    * `.Rmd` for documents
    * `.csv` for data
- Version control all these files with Git
- Ensure full reproducibility of the R environment with the `renv` package, which takes a snapshot of all packages used (and their version numbers), and can restore this configuration on a new system
- Publish the entire self-contained research unit on GitHub
- Connect this GitHub repository to an OSF project

Using text-based files has the advantage that they can be read by machines and humans without requiring special (licensed) software. Text-based files are also small, can be version-controlled with Git, and can be readily uploaded to online repositories.

-->

@@

TOP encourages comprehensive citation of literature, data, materials, methods, and software. In principle, researchers can meet this requirement by simply citing every reference used. In practice, however, print journals often discourage comprehensive citation due to space constraints. <!--either actively, or passively by including the reference list in the manuscript word count.--> Researchers can nonetheless meet the  comprehensive citation requirement by preparing two versions of the manuscript: A version with comprehensive citations for online dissemination (e.g., through preprint servers), and a print version that includes only essential citations. The Open Science statement in the print version should address the fact that additional references are available in the online version, e.g.:

> In the spirit of Open Science, an online repository is available for this manuscript at XXX, which contains more comprehensive citations, the data, analysis code, new research materials, and pre-registration forms.

The `worcs` package includes functions that can be used render a manuscript with, or without, comprehensive citations. The procedure to use these functions is:

1. During writing, maintain a plain-text `.bib` file with the BibTeX references for all citations.
  + You can make this file by hand; e.g., Figure \@ref(fig:scholarbib) shows how to obtain a BibTeX reference from Google Scholar; simply copy-paste each reference into the `.bib` file
  + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly. [Here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/) is a tutorial for using Zotero with Rmarkdown.
2. To cite a reference, use the `citekey` - the first word in the BibTeX entry for that reference. Insert it in the Rmarkdown file like so: `@yourcitekey2020`. For a parenthesized reference, use `[@citekeyone2020; @citekeytwo2020]`. For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
4. When Knitting the document, adapt the `knit` command in the YAML header.  
  `knit: worcs::cite_all` renders all citations, and  
  `knit: worcs::cite_essential` removes all *non-essential* citations.
5. Optional: To be extremely thorough, you could make a "branch" of the Github repository for the print version of the manuscript. Only in this branch, you use the function `knit: worcs::cite_essential`. The procedure is documented in [this tutorial](http://rstudio-pubs-static.s3.amazonaws.com/142364_3b344a38149b465c8ebc9a8cd2eee3aa.html).

@@

<!--1) use comprehensive citation, 2) document which citations are essential, and 3) publishing a pre-print with comprehensive citations, and 4) filtering out non-essential citations for the print version, and referencing the repository with the pre-print.

To facilitate the citation of data, researchers can publish a data paper in specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), or publish their data in an online repository - such as a GitHub repository - and include a text file with the preferred citation for the data, and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/).

Full version that includes all citations
Branch for publication version, with suppressed non-essential citations

Explanation for citations? How best to include
* Note
* In a function
* In a separate document-->

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, expliore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. In some cases, it is difficult or impossible to share data.

When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applying them to the raw data file. This way, the code will be a pipeline from the raw data to the published results.

@@

The second solution is to share a synthetic dataset with similar characteristics to the real data. Synthetic data typically mimic the level of measurement and (conditional) distributions of the real data, but complex multivariate relationships are often lost. Thus, analyses on the synthetic data should not be substantively interpreted, and findings from the real data might not be replicated in the synthetic data. Still, sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses. A third party can send this code to the original authors, who evaluate it on the real data, and send back the results. The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.

The `worcs` package includes three convenience functions to facilitate working with open and closed data. These functions are:

* `open_data(data)`: This function on a data object (a `data.frame` or `matrix`) generates a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`.
* `closed_data(data)`: This function generates a file called `data.csv` and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`.
* `load_data()`: This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data.

The WORCS-template additionally creates a file called `prepare_data.R`. This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses. Use this file to report any preprocessing required to obtain a neat `data.frame`. For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file. Typically, the last line of the `prepare_data.R` file will be `open_data()` or `closed_data()`. The proper analysis code will then start with `load_data()`.

@@

(3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis

@@

Which studies can be preregistered is a matter of some debate. Some have taken the position that studies can only be preregistered prior to data collection. This is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known. From this point of view, it makes little

@@

1. Document study plans in a `.Rmd` file, and planned analyses in a `.R` file.
2. Commit these documents to the local Git repository, and push them to GitHub
3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository

@@

WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-4), and preregistration (6-7).

@@

1. Create a new Private repository on github, copy the https:// link to clipboard  
  The link should look something like https://github.com/yourname/yourrepo.git
2. In Rstudio, click File > New Project > New directory > WORCS Project Template
    a. Paste the GitHub Repository address in the textbox
    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended)
    c. Select a preregistration template
3. Write the preregistration `.Rmd`
4. In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button. Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
7. Optional: Add study Materials (to which you own the rights) to the repository. It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)

@@

13. To save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
14. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF

@@

13. Make the GitHub repository public
14. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
15. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
16. Add an Open Science statement to the manuscript, with a link to the OSF project
17. Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.

<!-- Now, the worcs package adds a template to Rstudio

Template repository 
Template repositories let users generate new repositories with the same directory structure and files. Indicate if cjvanlissa/tidySEM can be used as a template for creating other repositories.

You’ve registered a free GitHub account (chapter 4).
You’ve installed/updated R and RStudio (chapter 5).
You’ve installed Git (chapter 6).
You’ve introduced yourself to Git (chapter 7).
You’ve confirmed that you can push to / pull from GitHub from the command line (chapter 9).

3. Write conceptual preregistration
* https://osf.io/zab38/wiki/home/ Templates of OSF Registration Forms
* R package prereg
* mention other examples and cite van 't Veer, A. E., & Giner-Sorolla, R. (2016). Pre-registration in social psychology---A discussion and suggested template. Journal of Experimental Social Psychology, 67, 2--12. doi: 10.1016/j.jesp.2016.03.004
3. Commit, push, and tag as preregistration
<!--Preregistration: tag on github
https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository

4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
-->

@@

* Dealing with collaborators
    - Add to Repo
    - Add Issues for tasks that still need to be accomplished, and tag Collaborators to address them
    - Handle pull requests from collaborators
* advantages
* limitations
* future directions

>>>>

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud, increasing awareness of questionable research practices, and the replication crisis. However, Open science is not merely a cure for this crisis - it is also an opportunity. Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science. Capitalizing on these advances has the potential to accellerate scientific progress.

This paper introduces WORCS: A Workflow for Open Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible. WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals. The main goal of WORCS is to lower the threshold for adopting an Open Science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.

<!--
thereby lowering the threshold for adoption.
, and still ensures reproducibility in most circumstances. 
 It is easy to implement, thereby lowering the threshold

* Grass-roots
* Open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction

Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
Collective benefits: Cumulative science, -->

@@

One of the most impactful contributions to Open Science has been the development of the TOP-guidelines, and the effort to urge publishers to adopt them. These guidelines describe eight standards for Open Science: 1) Comprehensive citation of literature, data, materials, and methods; sharing 2) data, 3) code required to reproduce analyses, 4) new research materials, and 5) details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results. WORCS defines the goals of Open Science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the topic of a research project, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations. Yet there is also substantial momentum for grassroots change. Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support. The formation of "Open Science Communities" at universities around the globe illustrates this movement. This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of Open Science. This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist, facilitates meeting them.

@@

@@

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science. Second, all of the aforementioned tools for an Open Science workflow are implemented in R. In general, many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Third, at the time of writing, R is the second-most cited statistical software package [@PopularityDataScience2012], following SPSS, which has no support for any of the tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.

Working with R is simplified immensely by using the Rstudio user interface for R. Rstudio integrates many of the tools used in WORCS in a user-friendly manner, and automates and streamlines many tedious or complicated aspects of working with R. This renders Rstudio a catch-all solution for generating open, reproducible code. One important feature of Rstudio is project management. A project bundles writing, analyses, data, references, etcetera, into a self-contained unit that can be uploaded to GitHub and downloaded by future users. The `worcs` R-package installs a new Rstudio project template, that performs much of the bookkeeping required to set up an Open Science project automatically  behind the scenes.

Another key consideration is what filetypes to use. A case can be made to use text-based files only (or as much as possible), instead of binary files. Text-based files can be read by machines and humans alike, without requiring special (licensed) software. Binary files must be decoded by software first. Git is designed to keep track of changes to the content of text-based files. Each change reflects a word or line of code added or removed. Binary files such as `.docx` or `.sav`, on the other hand, should not be version controlled. The changes in the encrypted file are uninterpretable. Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon. GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes. For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.

@@

- Use RStudio projects to bundle analyses and writing into a self-contained unit
- Use text-based files only
    * `.R` for analyses
    * `.Rmd` for documents
    * `.csv` for data
- Version control all these files with Git
- Ensure full reproducibility of the R environment with the `renv` package, which takes a snapshot of all packages used (and their version numbers), and can restore this configuration on a new system
- Publish the entire self-contained research unit on GitHub
- Connect this GitHub repository to an OSF project
-->

@@

TOP encourages comprehensive citation of literature, data, materials, methods, and software. In principle, researchers can meet this requirement by simply citing every reference used. Unfortunately, citation of data and software is less commonplace than citation of literature and materials. Crediting these resources is important, because it incentivizes data sharing and the development of open-source software, and supports the open science efforts of others.

<!--Various initiatives are underway to innovate how scientific output is measured and valued. https://www.nwo.nl/beleid/statement+waarderen+en+belonen+van+wetenschappers --> 
To facilitate citing datasets, researchers sometimes publish *data papers*; documents that detail the procedure, sample, and codebook. Specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), aid in the publication of these data papers. For smaller projects, researchers sometimes publish their data in an online repository - like GitHub - along with a text file with the preferred citation for the data (which can be a substantive paper), and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt, one can always contact the data creators and ask what the preferred citation is.

When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`. This returns an APA-style reference, and a BibTeX entry. However, users of the WORCS project template for Rstudio do not need to manually cite R-packages. WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.

One important impediment to comprehensive citation is the fact that print journals operate with space constraints. Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count. Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations. The print version should reference the online version in the print version, so interested readers can find the comprehensive reference list. Step 20 of the WORCS procedure suggests uploading the online version to a preprint server. This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar. This means that authors will receive credit for cited work; even if they are cited only in the online version. Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference. Users can render a manuscript to PDF either with, or without, comprehensive citations.

The procedure for citation in WORCS is as follows:

1. During writing, maintain a plain-text `.bib` file with the BibTeX references for all citations.
  + You can make this file by hand; e.g., Figure \@ref(fig:scholarbib) shows how to obtain a BibTeX reference from Google Scholar; simply copy-paste each reference into the `.bib` file
  + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly, and highly interoperable with other commercial reference managers. [Here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/) is a tutorial for using Zotero with Rmarkdown.
2. To cite a reference, use the `citekey` - the first word in the BibTeX entry for that reference. Insert it in the Rmarkdown file like so: `@yourcitekey2020`. For a parenthesized reference, use `[@citekeyone2020; @citekeytwo2020]`. For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
4. When Knitting the document, adapt the `knit` command in the YAML header.  
  `knit: worcs::cite_all` renders all citations, and  
  `knit: worcs::cite_essential` removes all *non-essential* citations.
5. Optional: To be extremely thorough, you could make a "branch" of the Github repository for the print version of the manuscript. Only in this branch, you use the function `knit: worcs::cite_essential`. The procedure is documented in [this tutorial](http://rstudio-pubs-static.s3.amazonaws.com/142364_3b344a38149b465c8ebc9a8cd2eee3aa.html).

@@

<!--
Future: Explain why each reference is cited? How best to include this?
* Note
* In a function
* In a separate document-->

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. In some cases, it is difficult or impossible to share data.

When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.

@@

The second solution is to share a synthetic dataset with similar characteristics to the real data. Synthetic data mimic the level of measurement and (conditional) distributions of the real data. Sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses. However, it is important to know that complex multivariate relationships present in the real data are often lost in synthetic data. Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*. Still, sharing synthetic data facilitates data requests from third parties. A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results. The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.

The procedure for data sharing in WORCS is as follows. First, researchers load the data into R. This process should be documented in the file `prepare_data.R`, which is automatically created by the WORCS-template. This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses. Use this file to report any preprocessing required to obtain a neat `data.frame`. For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file.

Then, researchers must decide whether to use open or closed data, and call one of these convenience functions:

* `open_data(data)`: This function takes a data object (a `data.frame` or `matrix`), writes it to a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`. The data can now be committed and pushed to GitHub, where they will be open once the GitHub repository is made public.
* `closed_data(data)`: This function generates a local file called `data.csv`, and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`. The synthetic data can now be committed and pushed to GitHub, and the real data will be ignored by Git.

Typically, the last line of the `prepare_data.R` file will thus be `open_data()` or `closed_data()`. The code for analysis of the data, on the other hand, will typically begin with `load_data()`. This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data. This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.

@@

When writing a manuscript in Rmarkdown, analysis code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing analysis code can be met simply by committing the source code of the manuscript to GitHub, and making this GitHub repository Public upon publication. If authors additionally use open data and a reproducible environment (default in WORCS), then a third party can simply replicate all analyses by copying the entire GitHub repository, and Knitting the manuscript on their local computer.

Aside from analysis code, the TOP-guidelines also encourage sharing new research materials, and details of the study design and analysis. These goals can be accomplished simply by placing any relevant documents in the Rstudio project folder, committing them to the Git repository, and pushing to GitHub. As with any document version controlled in this way, it is advisable to use plain text only.

@@

<!--Which studies can be preregistered is a matter of some debate. Some have taken the position that studies can only be preregistered prior to data collection. This is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known. From this point of view, it makes little -->

1. Document study plans in a `preregistration.Rmd` file, and planned analyses in a `.R` file.
2. Commit these documents to the local Git repository, and push them to GitHub
3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository

@@

@@

@@

1. Create a new Private repository on github, copy the https:// link to clipboard  
  The link should look something like https://github.com/yourname/yourrepo.git
2. In Rstudio, click File > New Project > New directory > WORCS Project Template
    a. Paste the GitHub Repository address in the textbox
    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended). The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
    c. Select a preregistration template
    d. Click Open in new session, and Create project
3. Write the preregistration `.Rmd`
4. In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button. Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
7. Optional: Add study Materials (to which you own the rights) to the repository. It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)

@@

13. Save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
14. Add an Open Science statement to the manuscript, which references the online repository for the project. This reference can be masked for blind review. The Open Science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available. For further guidance, see @aalbersbergMakingScienceTransparent2018. Example:  
    > In the spirit of Open Science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.

15. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF

@@

16. Make the GitHub repository public
17. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
19. Add an Open Science statement to the manuscript, with a link to the OSF project
20. Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.

<!-- 
4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
-->

@@

In this tutorial paper, we have presented a workflow for open reproducible code in science. The workflow aims to lower the threshold for grass-roots adoption of open science principles. It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio. This relatively light-weight workflow meets most of the requirements for Open Science, as detailed in the TOP-guidelines. However, several limitations and issues for future development remain.

One important remaining issue is how to deal with collaborators when using WORCS. Several different scenarios exist. In one scenario, only the lead author uses WORCS. In this case, it is possible to Knit the manuscript to Microsoft Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents). When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word. Then, manually incorporate their changes into the `manuscript.Rmd` file. In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.

A second approach is to send the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file. When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file. In Rstudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.

A third scenario is when all collaborators are committed to using WORCS. In this case, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes. Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013]. Moreover, being familiar with this workflow opens doors to new forms of collaboration. For example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality.

Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users. This package enables bidirectional conversion between Word and Rmarkdown, with support for tracked changes. However, a simpler solution (i.e., asking collaborators to work in plain text) might be sufficient.

    - Add to Repo
    - Add Issues for tasks that still need to be accomplished, and tag Collaborators to address them
    - Handle pull requests from collaborators
* advantages
* limitations
* future directions
